Dynamic Walking and Whole-Body Motion Planning for Humanoid Robots: an Integrated Approach

### 2012-12-06 paper.tex
Proposal to take into account jpl comments.
  * Comment about constrained RRT (un)usefulness not addressed yet.

\textbf{In this paper, we present a planning algorithm that considers
  exact models of a humanoid robot and its environment. It is used to
  solve navigation and manipulation problems. Our planner generates
  collision-free, statically balanced geometric paths, i.e. paths such
  that for any configuration along the path, the robot is balanced on
  its feet with zero speed. We also present a constructive formal
  proof that these paths can be approximated by collision-free,
  dynamically balanced walk trajectories, i.e. trajectories satisfying
  the Newton-Euler equations. Our planner is designed for perfectly
  modeled indoor environments, where the floor is horizontal and
  flat. Also, because we do not compute explicitly footstep positions
  at the planning stage, the presented planner is unable to plan
  motions in which the robot steps over obstacles. These limitations
  are discussed in the paper.}

@@

\textbf{In this work, we show that, under some assumptions, it is
  sufficient to plan in the configuration space in a first step, even
  though the final output of our algorithm is guaranteed to be a
  dynamically balanced walk motion. This result was first presented in
  \cite{dalibard2011small}.}

@@

Another contribution of this paper is the formal proof that this first
path can always be approximated by a dynamically balanced,
collision-free walking trajectory. We have implemented this
well-grounded method, and the results have been validated on the HRP-2
robot.

@@

\textbf{This section presents an algorithm for constrained motion
  planning in the configuration space $\mathcal{C}$. In the particular
  case of a humanoid robot which is an under-actuated system,
  $\mathcal{C}=\mathcal{Q} \times SE(3)$, where $\mathcal{Q}$
  represents the kinematic tree actuators and $SE(3)$ represents the
  position in space of the root of the tree. If the robot has $n$
  actuated DoFs, then $dim(\mathcal{Q})=n$ and
  $dim(\mathcal{C})=n+6$. A configuration $q$ of $\mathcal{C}$ is said
  to be valid iff, besides being collision-free, it lies on a specific
  submanifold $\manifold$ of $\mathcal{C}$. $\manifold$ is called the
  planning manifold.}

@@

If we use this algorithm with static balance constraints without
fixing globally the robot foot positions, it generates statically
balanced paths for a robot \textit{sliding} on the
ground. \textbf{Fig~\ref{fig:sliding} shows an example of a whole-body
  collision-free path for a robot passing between two chairs. Since in
  reality a legged robot cannot slide on a regular floor, such paths
  are physically unfeasible.They are, however, easier to generate than
  feasible dynamic walking trajectories because only geometric
  constraints are considered at planning time.}

>>>>

\textbf{In this paper, we present a planning algorithm that considers
  exact models of a humanoid robot and its environment. It is used to
  solve navigation and manipulation problems. Our planner is a
  two-step algorithm: a first collision-free path is computed in the
  quasi-static configuration space \footnote{A quasi-static
    configuration is a configuration where the robot is balanced on
    its feet with zero speed.}, then this first path is approximated
  by a sequence of dynamic walking trajectories. The proof of
  correctness of the algorithm is based on the concept of small-space
  controllability. Such property allows to approximate any \textit{non
    necessarily admissible} path, by a sequence of admissible ones. In
  our context we prove that dynamic walking makes humanoid robots
  small-space controllable. Our planner is designed for perfectly
  modeled indoor environments, where the floor is horizontal and
  flat. Also, because we do not explicitly compute footprint positions
  at the planning stage, the presented planner is unable to plan
  motions in which the robot steps over obstacles. These limitations
  are discussed in the paper.}

@@

\textbf{We show in this work that, under some assumptions, it is
  sufficient to plan a first path in the quasi-static configuration
  space of a humanoid robot, and then use the small-space
  controllability property to approximate this path by an admissible
  trajectory, i.e. a dynamically balanced walking motion. This result
  was first presented in \cite{dalibard2011small}.}

@@

Another contribution of this paper is the formal proof that ynamic
walking makes humanoid robots small-space controllable, whith the
direct implication that this first path can always be approximated by
a dynamically balanced, collision-free walking trajectory. We have
implemented this well-grounded method, and the results have been
validated on the HRP-2 robot.

@@

\textbf{This section presents an algorithm for constrained motion
  planning on a submanifold $\manifold$ of the configuration space
  $\mathcal{C}$. In the particular case of a humanoid robot which is
  an under-actuated system, $\mathcal{C}=\mathcal{Q} \times SE(3)$,
  where $\mathcal{Q}$ represents the kinematic tree actuators and
  $SE(3)$ represents the position in space of the root of the tree. If
  the robot has $n$ actuated DoFs, then $dim(\mathcal{Q})=n$ and
  $dim(\mathcal{C})=n+6$. A configuration $q$ of $\mathcal{C}$ is said
  to be valid iff, besides being collision-free, it lies on the
  manifold $\manifold$; we call $\manifold$ the planning manifold.}

@@

If we use this algorithm with static balance constraints without
fixing globally the robot foot positions, it generates statically
balanced paths for a robot \textit{sliding} on the
ground. \textbf{Fig~\ref{fig:sliding} shows an example of a whole-body
  collision-free path for a robot passing between two chairs. Since in
  reality a legged robot cannot slide on a regular floor, such paths
  are physically unfeasible. They are, however, easier to generate
  than feasible dynamic walking trajectories because only geometric
  constraints are considered at planning time.}
